Symbolic Dynamic Programming for Continuous State and Observation POMDPs

### 2012-10-11 cpomdp/camdp-xadd/doc/NIPS12/draft6.tex
File reorganization
Edits made on camera-ready

\author{
%Zahra Zamani \\
%ANU & NICTA\\
%Canberra, Australia \\
%\texttt{zahra.zamani@anu.edu.au} \\
%\And
%Coauthor \\
%Affiliation \\
%Address \\
%\texttt{email} \\
%\AND
%Coauthor \\
%Affiliation \\
%Address \\
%\texttt{email} \\
%\And
%Coauthor \\
%Affiliation \\
%Address \\
%\texttt{email} \\
%\And
%Coauthor \\
%Affiliation \\
%Address \\
%\texttt{email} \\
%(if needed)\\
}

@@

We use a dynamic Bayesian network (DBN) to compactly represent the transition model over the factored state variables and we likewise use a Bayesian network to represent the observation model:
{\footnotesize
\begin{align*}
\mathcal{T}: \;\; &
P(\vec{d_s}',\vec{x_s}'|\vec{d_s},\vec{x_s},a) = 
\prod_{i=1}^n P(d_{s_i}'|\vec{d_s},\vec{x_s},a) \prod_{j=1}^m P(x_{s_j}'|\vec{d_s},\vec{d_s}',\vec{x_s},a). \nonumber \\
\mathcal{Z}: \;\; & 
P(\vec{d_o},\vec{x_o}|\vec{d_s},\vec{x_s},a) = 
\prod_{i=1}^n P(d_{o_i}|d_{s_i},a) \prod_{j=1}^m P(x_{o_j}|x_{s_j},a). \nonumber 
\end{align*}}
For simplicity of algorithmic exposition, we do not allow synchronic
arcs in the transition representation, but note that to incorporate
such arcs only requires restrictions on the variable elimination
ordering during dynamic programming.  In these transition and
observation Bayes nets, 
probabilities over 
\emph{discrete} variables are represented using conditional probability
functions (CPF) in the form of
$P(d_i'|\vec{d},\vec{x},a)$. Probabilities for \emph{continuous} variables
$P(x_j'|\vec{d},\vec{d'},\vec{x},a)$ are represented by
\emph{piecewise linear functions} (PLFs) that are first-order Markov
and deterministic, i.e. the probabilities are encoded using the Dirac
$\delta[\cdot]$ function. The reward function $R(\vec{d},\vec{x},a)$
is set to a piecewise constant function. We next provide an intuitive
example to make this DC-POMDP framework more concrete.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Figure 1 - policy tree
\begin{figure}[t!]
%\vspace{-1mm}
\begin{center}
\includegraphics[width=0.4\textwidth]{pics/cond_plan2.pdf}
\hspace{5mm}
\includegraphics[width=0.45\textwidth]{pics/b1-4.pdf}
\end{center}
\vspace{-2mm}
\caption{\footnotesize (left) Example conditional plan $\beta^h$ for discrete observations; (right) an example piecewise $\alpha$-vector for continuous state $t$ represented
as a 
%The maximum $\alpha$ vector for a belief state at iteration 4 for the \textsc{\bf Power Plant} problem in Section~\ref{}
decision diagram: 
the \emph{true} branch is solid, the \emph{false}
branch is dashed.}
\label{fig:cond_plan}
\vspace{-1mm}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\begin{figure}[t]
%\begin{center}
%\includegraphics[width=0.4\textwidth]{pics/cond_plan2.pdf}
%\end{center}
%\vspace{-3mm}
%\caption{%\footnotesize 
%The optimal value function for \textsc{\bf Power Plant}
%as a decision diagram: 
%the \emph{true} branch is solid, the \emph{false}
%branch is dashed.} 
%\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\Begin
{
	$Z:=0$ \;	
		\For {all $\alpha_j$ in  $\Gamma^{h-1}$}    
		{
		\emph{// Point-based backup operation}\\
    	$\alpha_j(xd_s,xd_o) := \int_{x_s'} \bigoplus_{d_s'} P(xd_{o}|xd'_{s},a) \cdot P(xd'_{s}| xd_{s},a)\cdot \alpha_j \hspace{1mm}dx_s'$\;
		}  
		\For {all $j$ in $\alpha_j(xd_s,xd_o)$}    
		{
		\emph{// Generate value of $\alpha$-vector for belief $b_i$ as a function of observations}\\
		$\delta_{j}(xd_o) := $ $\int_{x_{s}} \bigoplus_{d_s} b_i \cdot \alpha_j(xd_s,xd_o)\hspace{1mm}dx_s$\\ \;
		}
		\For {all $j$ in $\delta_{j}(xd_o)$}    
		{
		\emph{// Generate partition points on the observation space}\\
		$Z := max(\delta_j(xd_o),z)$\;
    	}
    	\emph{// $\phi_{z_k}$ logical term on each partition of $Z$}\\
    	\emph{// Generate probabilities and partitions of observations that distinguish optimal $\alpha$-vector (conditional plan)}\\
   		$P(Z=z_k|b_i) := \int_{x_s'} \int_{x_o} \int_{x_s} \bigoplus_{d_o} \bigoplus_{d_s} \bigoplus_{d_s'} P(xd_o|xd'_s,a)*P(xd'_s|xd_s,a)*b_i* \mathbb{I}[\phi_{z_k}] d_{xd_o} d_{xd_s}d_{xd_s'}$ \;
    \Return{$P(z)$} \;
    %do this for each belief in B
}
\caption{\footnotesize \texttt{GenRelObs}($\Gamma^h,a,b_i$) $\longrightarrow$ $P(z)$ }
\label{alg:genrelobs}
\vspace{-1mm}
\end{algorithm}
\decmargin{.5em}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{PBVI for DC State and DC Observations} 
\label{sec:cont_obs}

@@

We use the second iteration of the \textsc{\bf Power Plant} example to
demonstrate partitioning of the entire continuous observation
space. The main assumption is to consider some belief state $b_i$ and
obtaining the relevant partitions according to that belief. For
example, we can choose two beliefs as uniform distributions,
e.g. $b_1: U[t_s;2,6]$ and $b_2: U[t_s;6,11]$ as shown in
Figure~\ref{fig:beliefs} (left).  Starting with an empty set of
relevant observation partitions $\mathcal{O}^h$, the algorithm first
eliminates the next state variable $xd'_s$. It regresses the
the set of
$\alpha$-vectors of the previous horizon back one step via
the transition and observation model: $\alpha_j(xd_s,xd_o) :=
\int_{x_s'} \bigoplus_{x_d'} P(xd_{o_i}|xd'_{s_i},a) * P(xd'_{s_i}| xd_{s_i},a)* \alpha_j\hspace{1mm}dx_s'
$.  This operation of line 5 in \texttt{GenRelObs} is similar to 
the point-based backup in~\eqref{eq:backup}. For
continuous variables the continuous integral and for discrete
variables the restriction operation is used as defined in
Section~\ref{sec:case}.  
Assuming that after $h=1$ the set of $\alpha$-vectors are
defined equal to the reward function of equation (1).
%{\footnotesize
%\vspace{-3mm}
%\begin{align}
%\alpha_1^1(t_s) &= 
%\begin{cases}
% (t_s>200) &: -1000 \\
%\neg(t_s>200) &: 100 \\
%\end{cases}
%\hspace{10mm} 
%\alpha_2^1(t_s) = \top: -1 \nonumber
%\end{align}
%}
For $h=2$ the resulting $\alpha$-vectors after the regression step are functions of the observation and current state: 
{\footnotesize
\vspace{-2mm}
\begin{align}
\alpha_1^2(t_s,t_o) &= 
\begin{cases}
 (t_s<15)\wedge (t_s - 10 < t_o<t_s) &: 10 \\
(t_s>15)\wedge (t_s - 10 < t_o<t_s) &: -100  \\
\neg(t_s - 10 < t_o<t_s) &: 0
\end{cases}
\hspace{2mm} 
\alpha_2^2(t_s,t_o) &= \begin{cases}
(t_s - 10 < t_o<t_s) &: -0.1 \\
\neg(t_s - 10 < t_o<t_s) &: 0
\end{cases}
\nonumber
\end{align}
} 
We now need the $\alpha$-vectors as a function of the observation
space for a particular belief state, thus next we eliminate the
current state variable in line 6--8. The resulting $\delta$-functions
now only depend on the observation: $\delta_{j}(xd_o) := $
$\int_{x_s'} \bigoplus_{x_d'} b_i * \alpha_j(xd_s,xd_o) \hspace{1mm}dx_s'$.  For $b_2: U[t_s;6,11]$
%assume the following simple $\delta$-functions:
the resulting $\delta$-functions are defined below:
{\footnotesize
%\vspace{-2mm}
\begin{align}
\delta_{1}(t_o) &= 
\begin{cases}
 (14 < t_o< 18) &: 0.025*t_o - 0.45\\
 (8 < t_o< 14) &:  - 0.1\\
 (4 < t_o< 8) &: - 0.025*t_o -0.1\\
\end{cases}
\hspace{5mm} 
\delta_{2}(t_o) &= \begin{cases}
 (15 < t_o< 18) &: 25*t_o - 450\\
 (14 < t_o< 15) &: - 2.5*t_o - 37.5\\
 (8 < t_o< 14) &:  -72.5\\
 (5 < t_o< 8) &: - 25*t_o + 127.5\\
 (4 < t_o< 5) &:  2.5*t_o - 10\\
\end{cases}
\nonumber
\end{align}
%\vspace{-5mm}
}
%{\footnotesize
%\begin{align}
%\delta_{1}(t_o) &= 
%\begin{cases}
% (2<t_o<6) &: 15 - 2.5*t_o \\
%(-4<t_o<2) &: 10 \\
%(-8<t_o<-4) &: 20 + 2.5*t_o 
%\end{cases}
%\hspace{5mm} 
%\delta_{2}(t_o) &= \begin{cases}
% (2<t_o<6) &: -0.15 + 0.025*t_o \\
%(-4<t_o<2) &: -0.1 \\
%(-8<t_o<-4) &: -0.2 - 0.025*t_o 
%\end{cases}
%\nonumber
%\end{align}
%}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}[tbp!]
\vspace{-3mm}
\centering
\hspace{-20mm}
\includegraphics[height=40mm]{pics/beliefs.pdf}
\hspace{-10mm}
\includegraphics[width=0.33\textwidth]{pics/delta_b1.pdf}
\hspace{-2mm}
\includegraphics[width=0.42\textwidth]{pics/delta_b2.pdf}
\hspace{-17mm}
\vspace{-2mm}
\caption{\footnotesize 
{\it (left)} Beliefs $b_1,b_2$ for 1D \textsc{\bf Power Plant} example; 
{\it (center)} Observation dependent function for $b_2$ that partition the observation space into 5 regions with different probabilities for $P(z_1),P(z_2)$ ; 
{\it (right)} 3 Relevant observation partitions (which is only one partition) and their probabilities for $b_2$.
}
\label{fig:beliefs}
%\vspace{-4mm}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
The observation dependent $\delta$-functions divide the observation space into regions which can yield the optimal policy according to the belief state $b_1$. According to continuous observation space defined in \cite{pascal_ijcai05}, for a 1-dimensional observation space with two or more functions of the observation, we need to find the optimal boundaries or partitions of the space. In their work, numerical solutions are proposed to find the roots of any two observation dependent function. Instead, here we leverage the symbolic power of the max-operator defined in Section~\ref{sec:case} to find all the boundary regions that define for what partitionings of the observation each 
$\delta$-function is optimal. For the two $\delta$-functions above the following partitions of the observation space is derived after taking their maximum in lines 9--11: 
{\footnotesize
\vspace{-2mm}
\begin{align}
\mathrm{max} \Bigg(\delta_{1}(t_o),\delta_{2}(t_o)\Bigg) &= 
\begin{cases}
z_1: (14 < t_o< 18) &: 0.025*t_o - 0.45\\
z_1: (8 < t_o< 14) &:  -0.1\\
z_1: (5.1 < t_o< 8) &: - 0.025*t_o -0.1\\
z_2: (5 < t_o< 5.1) &: - 25*t_o + 127.5\\
z_2: (4 < t_o< 5) &:  2.5*t_o - 10\\
\end{cases}
\nonumber
\end{align}
}
Note here that each partitioned is labeled according to the original $\delta$-function where $z_1$ is the label of the partition coming from $\delta_{1}(t_o)$. 
These partitions define observation regions which we can now use in a similar fashion to a discrete observation set if only the probability of each of the two distinct observation partitions is found.  This is demonstrated visually
in Figure~\ref{fig:beliefs} (center) for $b_2$ and for $b_1$
in Figure~\ref{fig:beliefs} (right). Note here that the maximum of the $\delta$-functions for $b_1$ is equal to $\delta_{2}(t_o)$.

Now with the observation partitions derived, all that remains is to
calculate probabilities for these relevant observations conditioned on
the belief state. 
% This does not check out!
% 
%As a reality check, we note that if we integrate out the
%state and observation from the observation model and the belief state
%regardless of a particular $\alpha$-vector, it integrates to one:
%$\int_{xd_o}\int_{xd_s} P(xd_o|xd'_s,a)*P(xd'_s|xd_s,a)*P(s) = 1$.
Thus we only need to multiply the indicators of each observation partition 
in this formula to obtain the probability mass lying in each
partition:
$$P(z_k|b_i) := \int_{xd_s'}\int_{xd_o}\int_{xd_s} P(xd_o|xd'_s,a)*P(xd'_s|xd_s,a)*b_i* \mathbb{I}[\phi_{z_k}] d_{xd_o} d_{xd_s}d_{xd_s'}$$
For our example the probabilities can be calculated as below:
{\footnotesize
\vspace{-2mm}
\begin{align}
\mathrm{max} \Bigg(\delta_{1}(t_o),\delta_{2}(t_o)\Bigg) &= 
\begin{cases}
(\phi_1 : 14 < t_o< 18) &: 0.2\\
(\phi_2 : 8 < t_o< 14) &:  0.6\\
(\phi_3 : 5.1 < t_o< 8) &: 0.183\\
(\phi_4 : 5 < t_o< 5.1) &: 0.002\\
(\phi_5 : 4 < t_o< 5) &:  0.0125\\
\end{cases}
\nonumber
\end{align}
}
where each constraint is defined as $\phi_i$. For any number of actions and observations the total number of observation partitions depends on the number of belief points. In our example we must have two observation probabilities which is defined as the sum of the probabilities for $z_1,z_2$ according to the constraints: \\
\begin{align*}
z_1&: \phi_1 \vee \phi_2 \vee \phi_3 \longrightarrow P(z_1|b_2)=0.983 \\
z_2&: \phi_4 \vee \phi_5  \longrightarrow P(z_2|b_2)=0.0127
\end{align*}

%{\footnotesize
%\begin{align}
%P(z_k)=
%\begin{cases}
% (t_o<2) &: 0.2 \\
%(2<t_o) &: 0.8\\
%\end{cases} 
%\nonumber
%\end{align}
%}

@@

In Figure ~\ref{fig:3D}  we present plots of the maximum $\alpha$-vectors of belief $b_1$ for different iterations of the 2D problem instance. %Each plot demonstrates the value of each vector as a function of the pressure and temperature. 
Starting with the first iteration, the value is highest for the reward range ($5<p<15 \wedge 95<t<105$) and -1 or less for other places. In the fifth iteration, the value function has partitioned into more pieces, showing how higher temperatures can increase the value without considering the effect of the pressure. In the last plot, horizon $h=6$ has better tuned the value function so that higher temperatures and pressures increase the value of the maximum $\alpha$-vector and also within the reward range, finer grain partitions have been formed. %Also note that the policy depends on the initial belief and for these plots they were based on $b_1$ which was lower than the reward range. The first action would then mean it has performed an increase while the other iterations decrease and increase respectively.  
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% policy2, v2plot.pdf, v9plot
% annotation?
\begin{figure*}[tbp!]
%\vspace{-2mm}
\centering
\includegraphics[width=0.31\textwidth]{pics/2d1.pdf}
\includegraphics[width=0.31\textwidth]{pics/2d9.pdf}
\includegraphics[width=0.31\textwidth]{pics/2d111.pdf}
\vspace{-3mm}
\caption{\footnotesize 
{\it (left)}Maximum $\alpha$-vector for $b_1$ in first iteration; 
{\it (center)} $\alpha_{max}^5(b_1)$; 
{\it (right)} $\alpha_{max}^5(b_1)$.%V^6(p,t)
}
\label{fig:3D}
%\vspace{-4mm}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%\subsubsection*{References} 
\bibliography{dcpomdp}
\bibliographystyle{plain}

>>>>

\author
{
Zahra Zamani \\
NICTA & ANU\\
Canberra, Australia \\
\texttt{zahra.zamani@anu.edu.au} \\
\And
Scott Sanner \\
NICTA & ANU\\
Canberra, Australia \\
\texttt{ssanner@nicta.com.au} \\
\And
Pascal Poupart \\
University of Waterloo \\
Waterloo, Canada \\
\texttt{ppoupart@uwaterloo.ca} \\
\And
Kristian Kersting \\
Fraunhofer IAIS, Sankt Augustin, Germany \\
IGG, University of Bonn, Germany \\
\texttt{kristian.kersting@iais.fraunhofer.de} \\
}

@@

We use a dynamic Bayesian network (DBN) to compactly represent the transition model over the factored state variables and we likewise use a Bayesian network to represent the observation model:
{\footnotesize
\begin{align*}
\mathcal{T}: \;\; &
P(\vec{d_s}',\vec{x_s}'|\vec{d_s},\vec{x_s},a) = 
\prod_{i=1}^n P(d_{s_i}'|\vec{d_s},\vec{x_s},a) \prod_{j=1}^m P(x_{s_j}'|\vec{d_s},\vec{d_s}',\vec{x_s},a). \nonumber \\
\mathcal{Z}: \;\; & 
P(\vec{d_o},\vec{x_o}|\vec{d_s},\vec{x_s},a) = 
\prod_{i=1}^n P(d_{o_i}|d_{s_i},a) \prod_{j=1}^m P(x_{o_j}|x_{s_j},a). \nonumber 
\end{align*}}
For simplicity of algorithmic exposition, we do not allow synchronic
arcs in the transition representation, but note that to incorporate
such arcs only requires restrictions on the variable elimination
ordering during dynamic programming.  In these transition and
observation Bayes nets, 
probabilities over 
\emph{discrete} variables are represented using conditional probability
functions (CPF) in the form of
$P(d_i'|\vec{d},\vec{x},a)$. Probabilities for \emph{continuous} variables
$P(x_j'|\vec{d},\vec{d'},\vec{x},a)$ are represented by
\emph{piecewise linear functions} (PLFs) that are first-order Markov
and deterministic, i.e. the probabilities are encoded using the Dirac
$\delta[\cdot]$ function. The reward function $R(\vec{d},\vec{x},a)$
is set to a piecewise constant function. We next provide an intuitive
example to make this DC-POMDP framework more concrete.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Figure 1 - policy tree
\begin{figure}[t!]
%\vspace{-1mm}
\begin{center}
\includegraphics[width=0.4\textwidth]{pics/cond_plan2.pdf}
\hspace{5mm}
\includegraphics[width=0.45\textwidth]{pics/dag2.pdf}
\end{center}
\vspace{-2mm}
\caption{\footnotesize (left) Example conditional plan $\beta^h$ for discrete observations; (right) an example piecewise $\alpha$-vector for continuous state $x$ represented
as a 
%The maximum $\alpha$ vector for a belief state at iteration 4 for the \textsc{\bf Power Plant} problem in Section~\ref{}
decision diagram: 
the \emph{true} branch is solid, the \emph{false}
branch is dashed.}
\label{fig:cond_plan}
\vspace{-1mm}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\begin{figure}[t]
%\begin{center}
%\includegraphics[width=0.4\textwidth]{pics/cond_plan2.pdf}
%\end{center}
%\vspace{-3mm}
%\caption{%\footnotesize 
%The optimal value function for \textsc{\bf Power Plant}
%as a decision diagram: 
%the \emph{true} branch is solid, the \emph{false}
%branch is dashed.} 
%\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\Begin
{
	$O:=0$ \;	
		\For {all $\alpha_j$ in  $\Gamma^{h-1}$}    
		{
		\emph{// Point-based backup operation}\\
    	$\alpha_j(xd_s,xd_o) := \int_{x_s'} \bigoplus_{d_s'} P(xd_{o}|xd'_{s},a) \cdot P(xd'_{s}| xd_{s},a)\cdot \alpha_j \hspace{1mm}dx_s'$\;
		}  
		\For {all $j$ in $\alpha_j(xd_s,xd_o)$}    
		{
		\emph{// Generate value of $\alpha$-vector for belief $b_i$ as a function of observations}\\
		$\delta_{j}(xd_o) := $ $\int_{x_{s}} \bigoplus_{d_s} b_i \cdot \alpha_j(xd_s,xd_o)\hspace{1mm}dx_s$\\ \;
		}
		\For {all $j$ in $\delta_{j}(xd_o)$}    
		{
		\emph{// Generate partition points on the observation space}\\
		$O := max(\delta_j(xd_o),\delta_{j+1}(xd_o))$\;
    	}
    	\emph{// $\phi_{o_k}$ logical term on each partition of $O$}\\
    	\emph{// Generate probabilities and partitions of observations that distinguish optimal $\alpha$-vector (conditional plan)}\\
   		$P(O=o_k|b_i) := \int_{x_s'} \int_{x_o} \int_{x_s} \bigoplus_{d_o} \bigoplus_{d_s} \bigoplus_{d_s'} P(xd_o|xd'_s,a)*P(xd'_s|xd_s,a)*b_i* \mathbb{I}[\phi_{o_k}] d_{x_o} d_{x_s}d_{x_s'}$ \;
    \Return{$P(o)$} \;
    %do this for each belief in B
}
\caption{\footnotesize \texttt{GenRelObs}($\Gamma^h,a,b_i$) $\longrightarrow$ $P(o)$ }
\label{alg:genrelobs}
\vspace{-1mm}
\end{algorithm}
\decmargin{.5em}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{PBVI for DC State and DC Observations} 
\label{sec:cont_obs}

@@

We use the second iteration of the \textsc{\bf Power Plant} example to
demonstrate partitioning of the entire continuous observation
space. The main assumption is to consider some belief state $b_i$ and
obtaining the relevant partitions according to that belief. For
example, we can choose two beliefs as uniform distributions,
e.g. $b_1: U[t_s;2,6]$ and $b_2: U[t_s;6,11]$ as shown in
Figure~\ref{fig:beliefs} (left).  Starting with an empty set of
relevant observation partitions $\mathcal{O}^h$, the algorithm first
eliminates the next state variable $xd'_s$. It regresses the
the set of
$\alpha$-vectors of the previous horizon back one step via
the transition and observation model: $\alpha_j(xd_s,xd_o) :=
\int_{x_s'} \bigoplus_{x_d'} P(xd_{o_i}|xd'_{s_i},a) * P(xd'_{s_i}| xd_{s_i},a)* \alpha_j\hspace{1mm}dx_s'
$.  This operation of line 5 in \texttt{GenRelObs} is similar to 
the point-based backup in~\eqref{eq:backup}. For
continuous variables the continuous integral and for discrete
variables the restriction operation is used as defined in
Section~\ref{sec:case}.  
Assuming that after $h=1$ the set of $\alpha$-vectors are
defined equal to the reward function of equation (1).
%\vspace{-3mm}
%\alpha_1^1(t_s) &= 
% (t_s>200) &: -1000 \\
%\neg(t_s>200) &: 100 \\
%\end{cases}
%\hspace{10mm} 
%\alpha_2^1(t_s) = \top: -1 \nonumber
For $h=2$ the resulting $\alpha$-vectors after the regression step are functions of the observation and current state: 
\alpha_{close}^2(t_s,t_o) &= 
 (t_s<15)\wedge (t_s - 10 < t_o<t_s) &: 10 \\
(t_s>15)\wedge (t_s - 10 < t_o<t_s) &: -100  \\
\neg(t_s - 10 < t_o<t_s) &: 0
\hspace{2mm} 
\alpha_{open}^2(t_s,t_o) &= \begin{cases}
(t_s - 10 < t_o<t_s) &: -0.1 \\
\neg(t_s - 10 < t_o<t_s) &: 0
\nonumber
We now need the $\alpha$-vectors as a function of the observation
space for a particular belief state, thus next we eliminate the
current state variable in line 6--8. The resulting $\delta$-functions
now only depend on the observation: $\delta_{j}^{b_i}(xd_o) := $
$\int_{x_s'} \bigoplus_{x_d'} b_i * \alpha_j(xd_s,xd_o) \hspace{1mm}dx_s'$.  For $b_2: U[t_s;6,11]$
%assume the following simple $\delta$-functions:
the resulting $\delta$-functions are defined below:
\delta_{close}^{b_2}(t_o) &= 
 (14 < t_o< 18) &: 0.025*t_o - 0.45\\
 (8 < t_o< 14) &:  - 0.1\\
 (4 < t_o< 8) &: - 0.025*t_o -0.1\\
\delta_{open}^{b_2}(t_o) &= \begin{cases}
 (15 < t_o< 18) &: 25*t_o - 450\\
 (14 < t_o< 15) &: - 2.5*t_o - 37.5\\
 (8 < t_o< 14) &:  -72.5\\
 (5 < t_o< 8) &: - 25*t_o + 127.5\\
 (4 < t_o< 5) &:  2.5*t_o - 10\\
\nonumber
%\vspace{-5mm}
%\delta_{1}(t_o) &= 
% (2<t_o<6) &: 15 - 2.5*t_o \\
%(-4<t_o<2) &: 10 \\
%(-8<t_o<-4) &: 20 + 2.5*t_o 
%\end{cases}
%\hspace{5mm} 
%\delta_{2}(t_o) &= \begin{cases}
% (2<t_o<6) &: -0.15 + 0.025*t_o \\
%(-4<t_o<2) &: -0.1 \\
%(-8<t_o<-4) &: -0.2 - 0.025*t_o 
%\end{cases}
%\nonumber
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}[tbp!]
\centering
\hspace{-20mm}
\includegraphics[width=0.42\textwidth]{pics/beliefs_2.pdf}
\hspace{10mm}
%\includegraphics[width=0.33\textwidth]{pics/delta_b1.pdf}
%\hspace{-2mm}
\includegraphics[width=0.42\textwidth]{pics/delta_b2_2.pdf}
\hspace{-17mm}
\caption{\footnotesize 
{\it (left)} Beliefs $b_1,b_2$ for 1D \textsc{\bf Power Plant} example; 
%{\it (center)} Observation dependent function for $b_2$ that partition the observation space into 5 regions with different probabilities for $P(o_1),P(o_2)$ ; 
{\it (right)} Relevant observation partitions and their probabilities for $b_2$. This diagrams shows that for temperatures $t_o < 5.1$ the policy is to open the valve and for higher temperatures, it is unsafe to open it and the policy should follow $\delta_{close}(t_o)$ which is to close the valve. 
\label{fig:beliefs}
%\vspace{-4mm}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
The observation dependent $\delta$-functions divide the observation space into regions which can yield the optimal policy according to the belief state $b_2$. According to continuous observation space defined in \cite{pascal_ijcai05}, for a 1-dimensional observation space with two or more functions of the observation, we need to find the optimal boundaries or partitions of the space. In their work, numerical solutions are proposed to find the roots of any two observation dependent function. Instead, here we leverage the symbolic power of the max-operator defined in Section~\ref{sec:case} to find all the boundary regions that define for what partitionings of the observation each 
$\delta$-function is optimal. For the two $\delta$-functions above the following partitions of the observation space is derived after taking their maximum in lines 9--11: 
\mathrm{max} \Bigg(\delta_{close}^{b_2}(t_o),\delta_{open}^{b_2}(t_o)\Bigg) &= 
o_1: (14 < t_o< 18) &: 0.025*t_o - 0.45\\
o_1: (8 < t_o< 14) &:  -0.1\\
o_1: (5.1 < t_o< 8) &: - 0.025*t_o -0.1\\
o_2: (5 < t_o< 5.1) &: - 25*t_o + 127.5\\
o_2: (4 < t_o< 5) &:  2.5*t_o - 10\\
\nonumber
Note here that each partitioned is labeled according to the original $\delta$-function where $o_1$ is the label of the partition coming from $\delta_{close}^{b_2}(t_o)$. 
These partitions define observation regions which we can now use in a similar fashion to a discrete observation set if only the probability of each of the two distinct observation partitions is found.  This is demonstrated visually
in Figure~\ref{fig:beliefs} (right) for $b_2$. %Note here that the maximum of the $\delta$-functions for $b_2$ is equal to $\delta_{open}^{b_2}(t_o)$.

Now with the observation partitions derived, all that remains is to
calculate probabilities for these relevant observations conditioned on
the belief state. 
% This does not check out!
% 
%As a reality check, we note that if we integrate out the
%state and observation from the observation model and the belief state
%regardless of a particular $\alpha$-vector, it integrates to one:
%$\int_{xd_o}\int_{xd_s} P(xd_o|xd'_s,a)*P(xd'_s|xd_s,a)*P(s) = 1$.
Thus we only need to multiply the indicators of each observation partition 
in this formula to obtain the probability mass lying in each
partition:
$$P(o_k|b_i) := \int_{x_s'}\int_{x_o}\int_{x_s} \bigoplus_{d_o} \bigoplus_{d_s} \bigoplus_{d_s'} P(xd_o|xd'_s,a)*P(xd'_s|xd_s,a)*b_i* \mathbb{I}[\phi_{o_k}] d_{x_o} d_{x_s}d_{x_s'}$$
For our example the non-zero probabilities occur in the following partitions as below:
{\footnotesize
\vspace{-2mm}
\begin{align}
P(o_k|b_2) &= 
\begin{cases}
(\phi_1 : 14 < t_o< 18) & : 0.2\\
(\phi_2 : 8 < t_o< 14) & :  0.6\\
(\phi_3 : 5.1 < t_o< 8) & : 0.183\\
(\phi_4 : 5 < t_o< 5.1) & : 0.0002\\
(\phi_5 : 4 < t_o< 5) & :  0.0125\\
\end{cases}
\nonumber
\end{align}
}
where each constraint is defined as $\phi_i$. For any number of actions and observations the total number of observation partitions depends on the number of belief points. In our example we must have two observation probabilities which is defined as the sum of the probabilities for $o_1,o_2$ according to the constraints: 
\begin{align*}
o_1&: \phi_1 \vee \phi_2 \vee \phi_3 \longrightarrow P(o_1|b_2)=0.983 \\
o_2&: \phi_4 \vee \phi_5  \longrightarrow P(o_2|b_2)=0.0127
\end{align*}
%{\footnotesize
%\begin{align}
%P(z_k)=
%\begin{cases}
% (t_o<2) &: 0.2 \\
%(2<t_o) &: 0.8\\
%\end{cases} 
%\nonumber
%\end{align}
%}

@@

In Figure ~\ref{fig:3D}  we present plots of the maximum $\delta$-vectors of belief $b_1$ for different iterations of the 2D problem instance. %Each plot demonstrates the value of each vector as a function of the pressure and temperature. 
Starting with the first iteration, the value is highest for the reward range ($5<p<15 \wedge 95<t<105$) and -1 or less for other places. In the fifth iteration, the value function has partitioned into more pieces, showing how higher temperatures can increase the value without considering the effect of the pressure. In the last plot, horizon $h=6$ has better tuned the value function so that higher temperatures and pressures increase the value of the maximum $\delta$-vector and also within the reward range, finer grain partitions have been formed. %Also note that the policy depends on the initial belief and for these plots they were based on $b_1$ which was lower than the reward range. The first action would then mean it has performed an increase while the other iterations decrease and increase respectively.  
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% policy2, v2plot.pdf, v9plot
% annotation?
\begin{figure*}[tbp!]
%\vspace{-2mm}
\centering
\includegraphics[width=0.31\textwidth]{pics/2d1-4.pdf}
\includegraphics[width=0.31\textwidth]{pics/2d9-4.pdf}
\includegraphics[width=0.31\textwidth]{pics/2d111-4.pdf}
\vspace{-3mm}
\caption{\footnotesize 
{\it (left)}Maximum $\delta$-vector for $b_1$ and action of $close$ in first iteration; 
{\it (center)} $\delta_{close}^5(b_1)$; 
{\it (right)} $\delta_{close}^6(b_1)$. Finer grain partitions show that closing (or opening) the valve can occur at more exact temperatures at higher horizons. 
}
\label{fig:3D}
%\vspace{-4mm}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\subsubsection*{Acknowledgments}

NICTA is funded by the Australian Government as represented by the Department of Broadband, Communications and the Digital Economy and the Australian Research Council through the ICT Centre of Excellence program. This work was supported by the Fraunhofer ATTRACT fellowship STREAM and by the EC, FP7-248258-First-MM.
%\subsubsection*{References} 
\bibliography{dcpomdp}
\bibliographystyle{plain}
